Forecasting the 2024 U.S. Presidential Election Results: Linear Regression, Bayesian and Spline-Fit Gaussian Model by Using Pollsters' Data From Five Thirty Eight

### 2024-11-04 paper/paper.qmd
update render pdf

library(tidyverse)
library(rstanarm)
library(splines)
library(broom)
library(knitr)
library(kableExtra)
```

@@

The remainder of this paper is structured as follows. @sec-data discusses the data used for this analysis, including key variables and sources, with particular attention to the quality metrics that affect polling accuracy. @sec-models outlines our modeling approach for each candidate, incorporating lessons learned from recent electoral cycles. @sec-predict presents our *** predictions based on the model outputs. @sec-discuss discusses the implications of our findings and suggests directions for future research. Finally, @sec-appendix evaluates Trafalgar Group’s polling methodology, and our idealized methodology and survey copy.

@@

Our data is download from [Five Thirty Eight] (https://projects.fivethirtyeight.com/polls) [@polls], the website gathered survey data from different pollsters of 2024 US president election, We use the statistical programming **language R** [@citeR], and packages **lubridate** [@lubridate], **dplyr** [@dplyr], **tidyverse** [@tidyverse], **model summary** [@models], and **arrow** [@arrow] to process the data. Also, we use **tibble** [@tibble] to create the table for simulate table and save it as csv files. Graphs are created by using ggplot2 in from package **tidyverse** [@tidyverse], packages **broom** [@broom] and **knitr** [@knitr] were used to generate clean summary and table in this paper. Following **Telling stories with data** [@telling_stories_with_data], we create different models for the candidates from different parties in general state and forecasting the US president election in 2024.

@@

# Create the density plot
ggplot(dem_data_filtered, aes(x = pct, fill = answer)) +
  geom_density(alpha = 0.4) +
  labs(
    x = "Support Percentage",
    y = "Density",
    fill = "Candidate"
  ) +
  theme_minimal()
```

>>>>

library(tidyverse)
library(rstanarm)
library(splines)
library(broom)
library(knitr)
library(patchwork)
library(kableExtra)
```

@@

The remainder of this paper is structured as follows. @sec-data discusses the data used for this analysis, including key variables and sources, with particular attention to the quality metrics that affect polling accuracy. @sec-models outlines our modeling approach for each candidate, incorporating lessons learned from recent electoral cycles. Our predictions are underr section of each model. @sec-discuss discusses the implications of our findings and suggests directions for future research. Finally, @sec-appendix evaluates Trafalgar Group’s polling methodology, and our idealized methodology and survey copy.

@@

Our data is download from [Five Thirty Eight] (https://projects.fivethirtyeight.com/polls) [@polls], the website gathered survey data from different pollsters of 2024 US president election, We use the statistical programming **language R** [@citeR], and packages **lubridate** [@lubridate], **dplyr** [@dplyr], **tidyverse** [@tidyverse], **model summary** [@models], and **arrow** [@arrow] to process the data. Also, we use **tibble** [@tibble] to create the table for simulate table and save it as csv files. Graphs are created by using ggplot2 in from package **tidyverse** [@tidyverse], packages **broom** [@broom] and **knitr** [@knitr] were used to generate clean summary and table in this paper. Package **patchwork** [@patchwork] is been used to combine two graphs together. Following **Telling stories with data** [@telling_stories_with_data], we create different models for the candidates from different parties in general state and forecasting the US president election in 2024.

@@

# Create the density plot
ggplot(dem_data_filtered, aes(x = pct, fill = answer)) +
  geom_density(alpha = 0.4) +
  labs(
    x = "Support Percentage",
    y = "Density",
    fill = "Candidate"
  ) +
  theme_minimal()
```
## Cross-Model Observations

In both the MLR and Bayesian models, Harris demonstrates a significant lead over Trump. The MLR model projects an approximate 2.6% lead for Harris, as indicated by the slope for the Democratic (DEM) variable, which is steeper and surpasses the Republican (REP) trend line between day 100 and day 200 (see @fig-mlrpredict). This early crossover point indicates a shift in support dynamics, suggesting that Harris’s policies or campaign events during this period may have positively influenced public sentiment. 
The Bayesian model, which accounts for random effects among pollsters, reinforces this lead by predicting 601 supporters for Harris versus 578 for Trump towards the end of the prediction period (see @fig-baypredict). This stability across models suggests that Harris’s support is likely durable, despite the model-specific variations in lead magnitude.
The Spline-Fit Gaussian model, however, presents a more conservative lead for Harris, with a smaller difference of approximately 1.06% in favor of the Democratic candidate by the end of the dataset (see @fig-splpredict). This model, designed to capture smoother trends, suggests that Harris’s lead, while present, may not be as robust as projected by the MLR and Bayesian models. This discrepancy emphasizes that different modeling approaches yield varied levels of confidence in support predictions, with the Spline-Fit model capturing subtle shifts that might indicate voter hesitancy or fluctuating preferences over time.
